find.interaction Find Interactions Between Pairs of Variables

#### Description
Find pairwise interactions between variables.
6 find.interaction

#### Usage 
<pre><code>
## S3 method for class 'rfsrc'
find.interaction(object, xvar.names, cause, m.target,
importance = c("permute", "random", "anti",
"permute.ensemble", "random.ensemble", "anti.ensemble"),
method = c("maxsubtree", "vimp"), sorted = TRUE, nvar, nrep = 1, subset,
na.action = c("na.omit", "na.impute"),
seed = NULL, do.trace = FALSE, verbose = TRUE, ...)
</code></pre> 
#### Arguments 
 
* ``object``: An object of class (rfsrc, grow) or (rfsrc, forest).
xvar.names Character vector of names of target x-variables. Default is to use all variables.
cause For competing risk families, integer value between 1 and J indicating the event
of interest, where J is the number of event types. The default is to use the first
event type.
m.target Character value for multivariate families specifying the target outcome to be
used. If left unspecified, the algorithm will choose a default target.
importance Type of variable importance (VIMP). See rfsrc for details.
method Method of analysis: maximal subtree or VIMP. See details below.
sorted Should variables be sorted by VIMP? Does not apply for competing risks.
nvar Number of variables to be used.
nrep Number of Monte Carlo replicates when ‘method="vimp"’.
subset Vector indicating which rows of the x-variable matrix from the object are to be
used. Uses all rows if not specified.
na.action Action to be taken if the data contains NA values. Applies only when ‘method="vimp"’.
seed Seed for random number generator. Must be a negative integer.
do.trace Number of seconds between updates to the user on approximate time to completion.
verbose Set to TRUE for verbose output.
... Further arguments passedto or from other methods.
Details
Using a previously grown forest, identify pairwise interactions for all pairs of variables from a
specified list. There are two distinct approaches specified by the option ‘method’.
1. ‘method="maxsubtree"’
This invokes a maximal subtree analysis. In this case, a matrix is returned where entries [i][i]
are the normalized minimal depth of variable [i] relative to the root node (normalized wrt the
size of the tree) and entries [i][j] indicate the normalized minimal depth of a variable [j] wrt the
maximal subtree for variable [i] (normalized wrt the size of [i]’s maximal subtree). Smaller
[i][i] entries indicate predictive variables. Small [i][j] entries having small [i][i] entries are a
sign of an interaction between variable i and j (note: the user should scan rows, not columns,
for small entries). See Ishwaran et al. (2010, 2011) for more details.
find.interaction 7
2. ‘method="vimp"’
This invokes a joint-VIMP approach. Two variables are paired and their paired VIMP calculated
(refered to as ’Paired’ importance). The VIMP for each separate variable is also calculated.
The sum of these two values is refered to as ’Additive’ importance. A large positive or
negative difference between ’Paired’ and ’Additive’ indicates an association worth pursuing if
the univariate VIMP for each of the paired-variables is reasonably large. See Ishwaran (2007)
for more details.
Computations might be slow depending upon the size of the data and the forest. In such cases,
consider setting ‘nvar’ to a smaller number. If ‘method="maxsubtree"’, consider using a smaller
number of trees in the original grow call.
If ‘nrep’ is greater than 1, the analysis is repeated nrep times and results averaged over the replications
(applies only when ‘method="vimp"’).
Value
Invisibly, the interaction table (a list for competing risk data) or the maximal subtree matrix.
Author(s)
Hemant Ishwaran and Udaya B. Kogalur

#### References

Ishwaran H. (2007). Variable importance in binary regression trees and forests, Electronic J. Statist.,
1:519-537.
* Ishwaran H., Kogalur U.B., Gorodeski E.Z, Minn A.J. and Lauer M.S. (2010). High-dimensional
variable selection for survival data. J. Amer. Statist. Assoc., 105:205-217.
* Ishwaran H., Kogalur U.B., Chen X. and Minn A.J. (2011). Random survival forests for highdimensional
data. Statist. Anal. Data Mining, 4:115-132.
See Also
max.subtree, var.select, vimp

#### Examples 
```{r}
## ------------------------------------------------------------
## find interactions, survival setting
## ------------------------------------------------------------
data(pbc, package = "randomForestSRC")
pbc.obj <- rfsrc(Surv(days,status) ~ ., pbc, importance = TRUE)
find.interaction(pbc.obj, method = "vimp", nvar = 8)
```
```{r}
## ------------------------------------------------------------
## find interactions, competing risks
## ------------------------------------------------------------

data(wihs, package = "randomForestSRC")
wihs.obj <- rfsrc(Surv(time, status) ~ ., wihs, nsplit = 3, ntree = 100,
importance = TRUE)
find.interaction(wihs.obj)
find.interaction(wihs.obj, method = "vimp")
```
```{r}
## ------------------------------------------------------------
## find interactions, regression setting
## ------------------------------------------------------------
airq.obj <- rfsrc(Ozone ~ ., data = airquality, importance = TRUE)
find.interaction(airq.obj, method = "vimp", nrep = 3)
find.interaction(airq.obj)
```
```{r}
## ------------------------------------------------------------
## find interactions, classification setting
## ------------------------------------------------------------
iris.obj <- rfsrc(Species ~., data = iris, importance = TRUE)
find.interaction(iris.obj, method = "vimp", nrep = 3)
find.interaction(iris.obj)
```
```{r}
## ------------------------------------------------------------
## interactions for multivariate mixed forests
## ------------------------------------------------------------
mtcars2 <- mtcars
mtcars2$cyl <- factor(mtcars2$cyl)
mtcars2$carb <- factor(mtcars2$carb, ordered = TRUE)
mv.obj <- rfsrc(cbind(carb, mpg, cyl) ~., data = mtcars2, importance = TRUE)
find.interaction(mv.obj, method = "vimp", outcome.target = "carb")
find.interaction(mv.obj, method = "vimp", outcome.target = "mpg")
find.interaction(mv.obj, method = "vimp", outcome.target = "cyl")
```


