impute Impute Only Mode

#### Description
Fast imputation mode. A random forest is grown and used to impute missing data. No ensemble
estimates or error rates are calculated.

#### Usage 
<pre><code>
## S3 method for class 'rfsrc'
impute(formula, data,
ntree = 500, nodesize = 1, nsplit = 10,
nimpute = 2, fast = FALSE, blocks,
mf.q, max.iter = 10, eps = 0.01,
ytry = NULL, always.use = NULL, verbose = TRUE,
...)
impute 11
</code></pre> 
#### Arguments 
 
formula A symbolic 
#### Description of the model to be fit. Can be left unspecified if there are
no outcomes or we don’t care to distinguish between y-outcomes and x-variables
in the imputation. Ignored when using multivariate missForest imputation.
data Data frame containing the data to be imputed.
ntree Number of trees to grow.
nodesize Forest average terminal node size.
nsplit Non-negative integer value used to specify random splitting.
nimpute Number of iterations of the missing data algorithm. Ignored for multivariate
missForest; in which case the algorithm iterates until a convergence criteria is
achieved (users can however enforce a maximum number of iterations with the
option max.iter).
fast Use fast random forests, rfsrcFast, in place of rfsrc? Improves speed but is
less accurate.
blocks Integer value specifying the number of blocks the data should be broken up into
(by rows). This can improve computational efficiency when the sample size is
large but imputation efficiency decreases. By default, no action is taken if left
unspecified.
mf.q Use this to turn on missForest (which is off by default). Specifies fraction of
variables (between 0 and 1) used as responses in multivariate missForest imputation.
Can also be an integer, in which case this equals the number of multivariate
responses.
max.iter Maximum number of iterations used when implementing multivariate missForest
imputation.
eps Tolerance value used to determine convergence of multivariate missForest imputation.
ytry Number of variables used as pseudo-responses in unsupervised forests. See
details below.
always.use Character vector of variable names to always be included as a response in multivariate
missForest imputation. Does not apply for other imputation methods.
verbose Send verbose output to terminal (only applies to multivariate missForest imputation).
... Further arguments passedto or from other methods.
Details
1. Grow a forest and use this to impute data. All external calculations such as ensemble calculations,
error rates, etc. are turned off. Use this function if your only interest is imputing the
data.
2. Split statistics are calculated using non-misssing data only. If a node splits on a variable with
missing data, the variable’s missing data is imputed by randomly drawing values from nonmissing
in-bag data. The purpose of this is to make it possible to assign cases to daughter
nodes based on the split.
12 impute
3. If no formula is specified, unsupervised splitting is implemented using a ytry value of sqrt(p)
where p equals the number of variables. More precisely, mtry variables are selected at random,
and for each of these a random subset of ytry variables are selected and defined as the
multivariate pseudo-responses. A multivariate composite splitting rule of dimension ytry is
then applied to each of the mtry multivariate regression problems and the node split on the
variable leading to the best split (Tang and Ishwaran, 2017).
4. If mf.q is specified, a multivariate version of missForest imputation (Stekhoven and Buhlmann,
2012) is applied. A fraction mf.q of variables are used as multivariate responses and split by
the remaining variables using multivariate composite splitting (Tang and Ishwaran, 2017).
Missing data for responses are imputed by prediction. The process is repeated using a new
set of variables for responses (mutually exclusive to the previous fit), until all variables have
been imputed. This is one iteration. The entire process is repeated, and the algorithm iterated
until a convergence criteria is met (specified using options max.iter and eps). Integer values
for mf.q are allowed and interpreted as a request that mf.q variables be selected for the multivariate
response. This is generally the most accurate of all the imputation procedures, but also
the most computationally demanding. However see 
#### Examples 
```{r} below for strategies to increase
speed.
5. Prior to imputation, the data is processed and records with all values missing are removed, as
are variables having all missing values.
6. If there is no missing data, either before or after processing of the data, the algorithm returns
the processed data and no imputation is performed.
7. The default choice nimpute=2 is chosen for coherence with the default missing data algorithm
implemented in grow mode. Thus, if the user imputes data with nimpute=2 and runs a grow
forest using this imputed data, then performance values such as VIMP and error rates will
coincide with those obtained by running a grow forest on the original non-imputed data using
na.action = "na.impute". Ignored for multivariate missForest.
8. All options are the same as rfsrc and the user should consult the rfsrc help file for details.
Value
Invisibly, the data frame containing the orginal data with imputed data overlayed.
Author(s)
Hemant Ishwaran and Udaya B. Kogalur

#### References

* Ishwaran H., Kogalur U.B., Blackstone E.H. and Lauer M.S. (2008). Random survival forests, Ann.
App. Statist., 2:841-860.
Stekhoven D.J. and Buhlmann P. (2012). MissForest–non-parametric missing value imputation for
mixed-type data. Bioinformatics, 28(1):112-118.
Tang F. and Ishwaran H. (2017). Random forest missing data algorithms. Statistical Analysis and
Data Mining, 10, 363-377.
See Also
rfsrc rfsrcFast
max.subtree 13

#### Examples 
```{r}
## ------------------------------------------------------------
## example of survival imputation
## ------------------------------------------------------------
## default everything - unsupervised splitting
data(pbc, package = "randomForestSRC")
pbc1.d <- impute(data = pbc)
## imputation using outcome splitting
f <- as.formula(Surv(days, status) ~ .)
pbc2.d <- impute(f, data = pbc, nsplit = 3)
## random splitting can be reasonably good
pbc3.d <- impute(f, data = pbc, splitrule = "random", nimpute = 5)
## ------------------------------------------------------------
## example of regression imputation
## ------------------------------------------------------------
air1.d <- impute(data = airquality, nimpute = 5)
air2.d <- impute(Ozone ~ ., data = airquality, nimpute = 5)
air3.d <- impute(Ozone ~ ., data = airquality, fast = TRUE)
## ------------------------------------------------------------
## multivariate missForest imputation
## ------------------------------------------------------------
data(pbc, package = "randomForestSRC")
## missForest algorithm - uses 1 variable at a time for the response
pbc.d <- impute(data = pbc, mf.q = 1)
## multivariate missForest - use 10 percent of variables as responses
## i.e. multivariate missForest
pbc.d <- impute(data = pbc, mf.q = .01)
## missForest but faster by using random splitting
pbc.d <- impute(data = pbc, mf.q = 1, splitrule = "random")
## missForest but faster by increasing nodesize
pbc.d <- impute(data = pbc, mf.q = 1, nodesize = 20, splitrule = "random")
## missForest but faster by using rfsrcFast
pbc.d <- impute(data = pbc, mf.q = 1, fast = TRUE)
