
partial Acquire Partial Effect of a Variable

#### Description
Acquire the partial effect of a variable on the ensembles.

#### Usage 
<pre><code>
partial.rfsrc(object, oob = TRUE, m.target = NULL,
partial.type = NULL, partial.xvar = NULL, partial.values = NULL,
partial.xvar2 = NULL, partial.values2 = NULL,
partial.time = NULL, seed = NULL, do.trace = FALSE, ...)
18 partial
</code></pre> 
#### Arguments 
 
* ``object``: An object of class (rfsrc, grow).
oob By default out-of-bag values are returned, but inbag values can be requested by
setting this option to FALSE.
m.target Character value for multivariate families specifying the target outcome to be
used. If left unspecified, the algorithm will choose a default target.
* ``partial.type``:  Character value of the type of predicted value. See details below.
* ``partial.xvar  ``:  Character value specifying the single primary partial x-variable to be used.
partial.values Vector of values that the primary partialy x-variable will assume.
* ``partial.xvar2 ``:  Vector of character values specifying the second order x-variables to be used.
partial.values2
Vector of values that the second order x-variables will assume. Each second order
x-variable can only assume a single value. This the length of partial.xvar2
and partial.values2 will be the same. In addition, the user must do the appropriate
conversion for factors, and represent a value as a numeric element.
* ``partial.time  ``:  For survival families, the time at which the predicted survival value is evaluated
at (depends on partial.type).
seed Negative integer specifying seed for the random number generator.
do.trace Number of seconds between updates to the user on approximate time to completion.
... Further arguments passedto or from other methods.
Details
Out-of-bag (OOB) values are returned by default.
A list of length equal to the number of outcomes (length is one for univariate families) with entries
depending on the underlying family:
1. For regression, the predicted response is returned of dim [n] x [length(partial.values)].
2. For classification, the predicted probabilities are returned of dim [n] x [1 + yvar.nlevels[.]] x [length(partial.values)].
3. For survival, the choices are:
• Relative frequency of mortality (rel.freq) or mortality (mort) is of dim [n] x [length(partial.values)].
• The cumulative hazard function (chf) is of dim [n] x [length(partial.time)] x
[length(partial.values)].
• The survival function (surv) is of dim [n] x [length(partial.time)] x [length(partial.values)].
4. For competing risks, the choices are:
• The expected number of life years lost (years.lost) is of dim [n] x [length(event.info$event.type)] x
[length(partial.values)].
• The cumulative incidence function (cif) is of dim [n] x [length(partial.time)] x
[length(event.info$event.type)] x [length(partial.values)].
• The cumulative hazard function (chf) is of dim [n] x [length(partial.time)] x [length(event.info$event.type)]
x [length(partial.values)].
partial 19
Author(s)
Hemant Ishwaran and Udaya B. Kogalur

#### References

* Ishwaran H., Kogalur U.B. (2007). Random survival forests for R, Rnews, 7(2):25-31.
* Ishwaran H., Kogalur U.B., Blackstone E.H. and Lauer M.S. (2008). Random survival forests, Ann.
App. Statist., 2:841-860.
See Also
plot.variable.rfsrc

#### Examples 
```{r}
## ------------------------------------------------------------
## survival/competing risk
## ------------------------------------------------------------
## survival
data(veteran, package = "randomForestSRC")
v.obj <- rfsrc(Surv(time,status)~., veteran, nsplit = 10, ntree = 100)
partial.obj <- partial(v.obj,
partial.type = "rel.freq",
partial.xvar = "age",
partial.values = v.obj$xvar[, "age"],
partial.time = v.obj$time.interest)
## competing risks
data(follic, package = "randomForestSRC")
follic.obj <- rfsrc(Surv(time, status) ~ ., follic, nsplit = 3, ntree = 100)
partial.obj <- partial(follic.obj,
partial.type = "cif",
partial.xvar = "age",
partial.values = follic.obj$xvar[, "age"],
partial.time = follic.obj$time.interest)
## regression
airq.obj <- rfsrc(Ozone ~ ., data = airquality)
partial.obj <- partial(airq.obj,
partial.xvar = "Wind",
partial.values = airq.obj$xvar[, "Wind"],
oob = FALSE)
## classification
iris.obj <- rfsrc(Species ~., data = iris)
partial.obj <- partial(iris.obj,
partial.xvar = "Sepal.Length",
partial.values = iris.obj$xvar[, "Sepal.Length"])
20 pbc
## multivariate mixed outcomes
mtcars2 <- mtcars
mtcars2$carb <- factor(mtcars2$carb)
mtcars2$cyl <- factor(mtcars2$cyl)
mtcars.mix <- rfsrc(Multivar(carb, mpg, cyl) ~ ., data = mtcars2)
partial.obj <- partial(mtcars.mix,
partial.xvar = "disp",
partial.values = mtcars.mix$xvar[, "disp"])
## second order variable specification
mtcars.obj <- rfsrc(mpg ~., data = mtcars)
partial.obj <- partial(mtcars.obj,
partial.xvar = "cyl",
partial.values = c(4, 8),
partial.xvar2 = c("gear", "disp", "carb"),
partial.values2 = c(4, 200, 3))

#### Examples 
```{r}
## ------------------------------------------------------------
## follicular cell lymphoma
## ------------------------------------------------------------
data(follic, package = "randomForestSRC")
follic.obj <- rfsrc(Surv(time, status) ~ ., follic, nsplit = 3, ntree = 100)
plot.competing.risk(follic.obj)
22 plot.rfsrc
## ------------------------------------------------------------
## competing risk analysis of pbc data from the survival package
## events are transplant (1) and death (2)
## ------------------------------------------------------------
if (library("survival", logical.return = TRUE)) {
data(pbc, package = "survival")
pbc$id <- NULL
plot.competing.risk(rfsrc(Surv(time, status) ~ ., pbc))
}
```
