## plot.variable Plot Marginal Effect of Variables

#### Description
Plot the marginal effect of an x-variable on the class probability (classification), response (regression),
mortality (survival), or the expected years lost (competing risk). Users can select between
marginal (unadjusted, but fast) and partial plots (adjusted, but slower).
Usage
## S3 method for class 'rfsrc'
plot.variable(x, xvar.names, target,
              m.target = NULL, time, surv.type = c("mort", "rel.freq",
                                                   "surv", "years.lost", "cif", "chf"), class.type =
                c("prob", "bayes"), partial = FALSE, oob = TRUE, show.plots = TRUE,
              plots.per.page = 4, granule = 5, sorted = TRUE, nvar, npts = 25,
              smooth.lines = FALSE, subset, ...)

#### Arguments
x An object of class (rfsrc, grow), (rfsrc, synthetic), or (rfsrc, plot.variable).
See the examples below for illustration of the latter.
* xvar.names Names of the x-variables to be used.
target For classification, an integer or character value specifying the class to focus on
(defaults to the first class). For competing risks, an integer value between 1 and
J indicating the event of interest, where J is the number of event types. The
default is to use the first event type.
* m.target Character value for multivariate families specifying the target outcome to be
used. If left unspecified, the algorithm will choose a default target.
time For survival, the time at which the predicted survival value is evaluated at (depends
                                                                                   on surv.type).
* surv.type For survival, specifies the predicted value. See details below.
* class.type For classification, specifies the predicted value. See details below.
partial Should partial plots be used?
  oob OOB (TRUE) or in-bag (FALSE) predicted values.
* show.plots Should plots be displayed?
  plots.per.page Integer value controlling page layout.
* granule Integer value controlling whether a plot for a specific variable should be treated
as a factor and therefore given as a boxplot. Larger values coerce boxplots.
sorted Should variables be sorted by importance values.
nvar Number of variables to be plotted. Default is all.
plot.variable 29
* npts Maximum number of points used when generating partial plots for continuous variables.
* smooth.lines Use lowess to smooth partial plots.
* subset Vector indicating which rows of the x-variable matrix x$xvar to use. All rows are used if not specified. Do not define subset based on the original data (which
                                                                            could have been processed due to missing values or for other reasons in the
                                                                            previous forest call) but define subset based on the rows of x$xvar.
... Further arguments passed to or from other methods.

#### Details
The vertical axis displays the ensemble predicted value, while x-variables are plotted on the horizontal
axis.
1. For regression, the predicted response is used.
2. For classification, it is the predicted class probability specified by ‘target’, or the class of
maximum probability depending on ‘class.type’.
3. For multivariate families, it is the predicted value of the outcome specified by ‘m.target’ and
if that is a classification outcome, by ‘target’.
4. For survival, the choices are:
  • Mortality (mort).
• Relative frequency of mortality (rel.freq).
• Predicted survival (surv), where the predicted survival is for the time point specified
using time (the default is the median follow up time).
5. For competing risks, the choices are:
  • The expected number of life years lost (years.lost).
• The cumulative incidence function (cif).
• The cumulative hazard function (chf).
In all three cases, the predicted value is for the event type specified by ‘target’. For cif and
chf the quantity is evaluated at the time point specified by time.
For partial plots use ‘partial=TRUE’. Their interpretation are different than marginal plots. The
y-value for a variable X, evaluated at X = x, is
$$
˜f(x) = 1 n Xn i=1 ˆf(x, xi,o),$$
where xi,o represents the value for all other variables other than X for individual i and ˆf is the
predicted value. Generating partial plots can be very slow. Choosing a small value for npts can
speed up computational times as this restricts the number of distinct x values used in computing ˜f.

* For continuous variables, red points are used to indicate partial values and dashed red lines indicate
a smoothed error bar of +/- two standard errors. Black dashed line are the partial values.
Set ‘smooth.lines=TRUE’ for lowess smoothed lines. For discrete variables, partial values are indicated
using boxplots with whiskers extending out approximately two standard errors from the
mean. Standard errors are meant only to be a guide and should be interpreted with caution.

* Partial plots can be slow. Setting ‘npts’ to a smaller number can help. For greater customization and flexibility in partial plot calls, consider using the function partial.rfsrc
which provides a direct interface for calculating partial plot data.

#### Author(s)
Hemant Ishwaran and Udaya B. Kogalur
#### References
Friedman J.H. (2001). Greedy function approximation: a gradient boosting machine, Ann. of
Statist., 5:1189-1232.
Ishwaran H., Kogalur U.B. (2007). Random survival forests for R, Rnews, 7(2):25-31.
Ishwaran H., Kogalur U.B., Blackstone E.H. and Lauer M.S. (2008). Random survival forests, Ann.
App. Statist., 2:841-860.
Ishwaran H., Gerds T.A., Kogalur U.B., Moore R.D., Gange S.J. and Lau B.M. (2014). Random
survival forests for competing risks. Biostatistics, 15(4):757-773.
See Also
rfsrc, rfsrcSyn, partial.rfsrc, predict.rfsrc
#### Examples
```{r}
## ------------------------------------------------------------
## survival/competing risk
## ------------------------------------------------------------
## survival
data(veteran, package = "randomForestSRC")
v.obj <- rfsrc(Surv(time,status)~., veteran, ntree = 100)
plot.variable(v.obj, plots.per.page = 3)
plot.variable(v.obj, plots.per.page = 2, xvar.names = c("trt", "karno", "age"))
plot.variable(v.obj, surv.type = "surv", nvar = 1, time = 200)
plot.variable(v.obj, surv.type = "surv", partial = TRUE, smooth.lines = TRUE)
plot.variable(v.obj, surv.type = "rel.freq", partial = TRUE, nvar = 2)
```
```{r}
## example of plot.variable calling a pre-processed plot.variable object
p.v <- plot.variable(v.obj, surv.type = "surv", partial = TRUE, smooth.lines = TRUE)
plot.variable(p.v)
p.v$plots.per.page <- 1
p.v$smooth.lines <- FALSE
plot.variable(p.v)
## competing risks
data(follic, package = "randomForestSRC")
follic.obj <- rfsrc(Surv(time, status) ~ ., follic, nsplit = 3, ntree = 100)
plot.variable(follic.obj, target = 2)
predict.rfsrc 31
```

```{r}
## ------------------------------------------------------------
## regression
## ------------------------------------------------------------
## airquality
airq.obj <- rfsrc(Ozone ~ ., data = airquality)
plot.variable(airq.obj, partial = TRUE, smooth.lines = TRUE)
plot.variable(airq.obj, partial = TRUE, subset = airq.obj$xvar$Solar.R < 200)
```


```{r}
## motor trend cars
mtcars.obj <- rfsrc(mpg ~ ., data = mtcars)
plot.variable(mtcars.obj, partial = TRUE, smooth.lines = TRUE)
```


```{r}
## ------------------------------------------------------------
## classification
## ------------------------------------------------------------
## iris
iris.obj <- rfsrc(Species ~., data = iris)
plot.variable(iris.obj, partial = TRUE)
## motor trend cars: predict number of carburetors
mtcars2 <- mtcars
mtcars2$carb <- factor(mtcars2$carb,
                       labels = paste("carb", sort(unique(mtcars$carb))))
mtcars2.obj <- rfsrc(carb ~ ., data = mtcars2)
plot.variable(mtcars2.obj, partial = TRUE)
```


```{r}
## ------------------------------------------------------------
## multivariate regression
## ------------------------------------------------------------
mtcars.mreg <- rfsrc(Multivar(mpg, cyl) ~., data = mtcars)
plot.variable(mtcars.mreg, m.target = "mpg", partial = TRUE, nvar = 1)
plot.variable(mtcars.mreg, m.target = "cyl", partial = TRUE, nvar = 1)
```


```{r}
## ------------------------------------------------------------
## multivariate mixed outcomes
## ------------------------------------------------------------
mtcars2 <- mtcars
mtcars2$carb <- factor(mtcars2$carb)
mtcars2$cyl <- factor(mtcars2$cyl)
mtcars.mix <- rfsrc(Multivar(carb, mpg, cyl) ~ ., data = mtcars2)
plot.variable(mtcars.mix, m.target = "cyl", target = "4", partial = TRUE, nvar = 1)
plot.variable(mtcars.mix, m.target = "cyl", target = 2, partial = TRUE, nvar = 1)

```
