plot.survival Plot of Survival Estimates
Description
Plot various survival estimates.
Usage
## S3 method for class 'rfsrc'
plot.survival(x, plots.one.page = TRUE,
              show.plots = TRUE, subset, collapse = FALSE,
              haz.model = c("spline", "ggamma", "nonpar", "none"),
              k = 25, span = "cv", cens.model = c("km", "rfsrc"), ...)
Arguments
x An object of class (rfsrc, grow) or (rfsrc, predict).
plots.one.page Should plots be placed on one page?
  show.plots Should plots be displayed?
  subset Vector indicating which individuals we want estimates for. All individuals are
used if not specified.
collapse Collapse the survival and cumulative hazard function across the individuals
specified by ‘subset’? Only applies when ‘subset’ is specified.
haz.model Method for estimating the hazard. See details below. Applies only when ‘subset’
is specified.
k The number of natural cubic spline knots used for estimating the hazard function.
Applies only when ‘subset’ is specified.
span The fraction of the observations in the span of Friedman’s super-smoother used
for estimating the hazard function. Applies only when ‘subset’ is specified.
26 plot.survival
cens.model Method for estimating the censoring distribution used in the inverse probability
of censoring weights (IPCW) for the Brier score:
  km: Uses the Kaplan-Meier estimator.
rfscr: Uses random survival forests.
... Further arguments passed to or from other methods.
Details
If ‘subset’ is not specified, generates the following three plots (going from top to bottom, left to
                                                                   right):
  1. Forest estimated survival function for each individual (thick red line is overall ensemble survival,
                                                             thick green line is Nelson-Aalen estimator).
2. Brier score (0=perfect, 1=poor, and 0.25=guessing) stratified by ensemble mortality. Based
on the IPCW method described in Gerds et al. (2006). Stratification is into 4 groups corresponding
to the 0-25, 25-50, 50-75 and 75-100 percentile values of mortality. Red line is the
overall (non-stratified) Brier score.
3. Plot of mortality of each individual versus observed time. Points in blue correspond to events,
black points are censored observations.
When ‘subset’ is specified, then for each individual in ‘subset’, the following three plots are
generated:
  1. Forest estimated survival function.
2. Forest estimated cumulative hazard function (CHF) (displayed using black lines). Blue lines
are the CHF from the estimated hazard function. See the next item.
3. A smoothed hazard function derived from the forest estimated CHF (or survival function).
The default method, ‘haz.model="spline"’, models the log CHF using natural cubic splines
as described in Royston and Parmar (2002). The lasso is used for model selection, implemented
using the glmnet package (this package must be installed for this option to work).
If ‘haz.model="ggamma"’, a three-parameter generalized gamma distribution (using the parameterization
                                                                           described in Cox et al, 2007) is fit to the smoothed forest survival function,
where smoothing is imposed using Friedman’s supersmoother (implemented by supsmu). If
‘haz.model="nonpar"’, Friedman’s supersmoother is applied to the forest estimated hazard
function (obtained by taking the crude derivative of the smoothed forest CHF). Finally, setting
‘haz.model="none"’ suppresses hazard estimation and no hazard estimate is provided.
At this time, please note that all hazard estimates are considered experimental and users should
interpret the results with caution.
Note that when the object x is of class (rfsrc, predict) not all plots will be produced. In
particular, Brier scores are not calculated.
Only applies to survival families. In particular, fails for competing risk analyses. Use plot.competing.risk
in such cases.
Whenever possible, out-of-bag (OOB) values are used.
plot.survival 27
Value
Invisibly, the conditional and unconditional Brier scores, and the integrated Brier score (if they are
                                                                                           available).
Author(s)
Hemant Ishwaran and Udaya B. Kogalur
References
Cox C., Chu, H., Schneider, M. F. and Munoz, A. (2007). Parametric survival analysis and taxonomy
of hazard functions for the generalized gamma distribution. Statistics in Medicine 26:4252-
  4374.
Gerds T.A and Schumacher M. (2006). Consistent estimation of the expected Brier score in general
survival models with right-censored event times, Biometrical J., 6:1029-1040.
Graf E., Schmoor C., Sauerbrei W. and Schumacher M. (1999). Assessment and comparison of
prognostic classification schemes for survival data, Statist. in Medicine, 18:2529-2545.
Ishwaran H. and Kogalur U.B. (2007). Random survival forests for R, Rnews, 7(2):25-31.
Royston P. and Parmar M.K.B. (2002). Flexible parametric proportional-hazards and proportionalodds
models for censored survival data, with application to prognostic modelling and estimation of
treatment effects, Statist. in Medicine, 21::2175-2197.
See Also
plot.competing.risk, predict.rfsrc, rfsrc
Examples
## veteran data
data(veteran, package = "randomForestSRC")
plot.survival(rfsrc(Surv(time, status)~ ., veteran), cens.model = "rfsrc")
## pbc data
data(pbc, package = "randomForestSRC")
pbc.obj <- rfsrc(Surv(days, status) ~ ., pbc)
# default spline approach
plot.survival(pbc.obj, subset = 3)
plot.survival(pbc.obj, subset = 3, k = 100)
# three-parameter generalized gamma is approximately the same
# but notice that its CHF estimate (blue line) is not as accurate
plot.survival(pbc.obj, subset = 3, haz.model = "ggamma")
# nonparametric method is too wiggly or undersmooths
plot.survival(pbc.obj, subset = 3, haz.model = "nonpar", span = 0.1)
plot.survival(pbc.obj, subset = 3, haz.model = "nonpar", span = 0.8)
28 plot.variable
plot.variable Plot Marginal Effect of Variables
Description
Plot the marginal effect of an x-variable on the class probability (classification), response (regression),
mortality (survival), or the expected years lost (competing risk). Users can select between
marginal (unadjusted, but fast) and partial plots (adjusted, but slower).
Usage
## S3 method for class 'rfsrc'
plot.variable(x, xvar.names, target,
              m.target = NULL, time, surv.type = c("mort", "rel.freq",
                                                   "surv", "years.lost", "cif", "chf"), class.type =
                c("prob", "bayes"), partial = FALSE, oob = TRUE, show.plots = TRUE,
              plots.per.page = 4, granule = 5, sorted = TRUE, nvar, npts = 25,
              smooth.lines = FALSE, subset, ...)
Arguments
x An object of class (rfsrc, grow), (rfsrc, synthetic), or (rfsrc, plot.variable).
See the examples below for illustration of the latter.
xvar.names Names of the x-variables to be used.
target For classification, an integer or character value specifying the class to focus on
(defaults to the first class). For competing risks, an integer value between 1 and
J indicating the event of interest, where J is the number of event types. The
default is to use the first event type.
m.target Character value for multivariate families specifying the target outcome to be
used. If left unspecified, the algorithm will choose a default target.
time For survival, the time at which the predicted survival value is evaluated at (depends
                                                                                   on surv.type).
surv.type For survival, specifies the predicted value. See details below.
class.type For classification, specifies the predicted value. See details below.
partial Should partial plots be used?
  oob OOB (TRUE) or in-bag (FALSE) predicted values.
show.plots Should plots be displayed?
  plots.per.page Integer value controlling page layout.
granule Integer value controlling whether a plot for a specific variable should be treated
as a factor and therefore given as a boxplot. Larger values coerce boxplots.
sorted Should variables be sorted by importance values.
nvar Number of variables to be plotted. Default is all.
plot.variable 29
npts Maximum number of points used when generating partial plots for continuous
variables.
smooth.lines Use lowess to smooth partial plots.
subset Vector indicating which rows of the x-variable matrix x$xvar to use. All rows
are used if not specified. Do not define subset based on the original data (which
                                                                            could have been processed due to missing values or for other reasons in the
                                                                            previous forest call) but define subset based on the rows of x$xvar.
... Further arguments passed to or from other methods.
Details
The vertical axis displays the ensemble predicted value, while x-variables are plotted on the horizontal
axis.
1. For regression, the predicted response is used.
2. For classification, it is the predicted class probability specified by ‘target’, or the class of
maximum probability depending on ‘class.type’.
3. For multivariate families, it is the predicted value of the outcome specified by ‘m.target’ and
if that is a classification outcome, by ‘target’.
4. For survival, the choices are:
  • Mortality (mort).
• Relative frequency of mortality (rel.freq).
• Predicted survival (surv), where the predicted survival is for the time point specified
using time (the default is the median follow up time).
5. For competing risks, the choices are:
  • The expected number of life years lost (years.lost).
• The cumulative incidence function (cif).
• The cumulative hazard function (chf).
In all three cases, the predicted value is for the event type specified by ‘target’. For cif and
chf the quantity is evaluated at the time point specified by time.
For partial plots use ‘partial=TRUE’. Their interpretation are different than marginal plots. The
y-value for a variable X, evaluated at X = x, is
˜f(x) = 1
n
Xn
i=1
ˆf(x, xi,o),
where xi,o represents the value for all other variables other than X for individual i and ˆf is the
predicted value. Generating partial plots can be very slow. Choosing a small value for npts can
speed up computational times as this restricts the number of distinct x values used in computing ˜f.
For continuous variables, red points are used to indicate partial values and dashed red lines indicate
a smoothed error bar of +/- two standard errors. Black dashed line are the partial values.
Set ‘smooth.lines=TRUE’ for lowess smoothed lines. For discrete variables, partial values are indicated
using boxplots with whiskers extending out approximately two standard errors from the
mean. Standard errors are meant only to be a guide and should be interpreted with caution.
30 plot.variable
Partial plots can be slow. Setting ‘npts’ to a smaller number can help.
For greater customization and flexibility in partial plot calls, consider using the function partial.rfsrc
which provides a direct interface for calculating partial plot data.
Author(s)
Hemant Ishwaran and Udaya B. Kogalur
References
Friedman J.H. (2001). Greedy function approximation: a gradient boosting machine, Ann. of
Statist., 5:1189-1232.
Ishwaran H., Kogalur U.B. (2007). Random survival forests for R, Rnews, 7(2):25-31.
Ishwaran H., Kogalur U.B., Blackstone E.H. and Lauer M.S. (2008). Random survival forests, Ann.
App. Statist., 2:841-860.
Ishwaran H., Gerds T.A., Kogalur U.B., Moore R.D., Gange S.J. and Lau B.M. (2014). Random
survival forests for competing risks. Biostatistics, 15(4):757-773.
See Also
rfsrc, rfsrcSyn, partial.rfsrc, predict.rfsrc
Examples
## ------------------------------------------------------------
## survival/competing risk
## ------------------------------------------------------------
## survival
data(veteran, package = "randomForestSRC")
v.obj <- rfsrc(Surv(time,status)~., veteran, ntree = 100)
plot.variable(v.obj, plots.per.page = 3)
plot.variable(v.obj, plots.per.page = 2, xvar.names = c("trt", "karno", "age"))
plot.variable(v.obj, surv.type = "surv", nvar = 1, time = 200)
plot.variable(v.obj, surv.type = "surv", partial = TRUE, smooth.lines = TRUE)
plot.variable(v.obj, surv.type = "rel.freq", partial = TRUE, nvar = 2)
## example of plot.variable calling a pre-processed plot.variable object
p.v <- plot.variable(v.obj, surv.type = "surv", partial = TRUE, smooth.lines = TRUE)
plot.variable(p.v)
p.v$plots.per.page <- 1
p.v$smooth.lines <- FALSE
plot.variable(p.v)
## competing risks
data(follic, package = "randomForestSRC")
follic.obj <- rfsrc(Surv(time, status) ~ ., follic, nsplit = 3, ntree = 100)
plot.variable(follic.obj, target = 2)
predict.rfsrc 31
```

```{r}
## ------------------------------------------------------------
## regression
## ------------------------------------------------------------
## airquality
airq.obj <- rfsrc(Ozone ~ ., data = airquality)
plot.variable(airq.obj, partial = TRUE, smooth.lines = TRUE)
plot.variable(airq.obj, partial = TRUE, subset = airq.obj$xvar$Solar.R < 200)
```


```{r}
## motor trend cars
mtcars.obj <- rfsrc(mpg ~ ., data = mtcars)
plot.variable(mtcars.obj, partial = TRUE, smooth.lines = TRUE)
```


```{r}
## ------------------------------------------------------------
## classification
## ------------------------------------------------------------
## iris
iris.obj <- rfsrc(Species ~., data = iris)
plot.variable(iris.obj, partial = TRUE)
## motor trend cars: predict number of carburetors
mtcars2 <- mtcars
mtcars2$carb <- factor(mtcars2$carb,
                       labels = paste("carb", sort(unique(mtcars$carb))))
mtcars2.obj <- rfsrc(carb ~ ., data = mtcars2)
plot.variable(mtcars2.obj, partial = TRUE)
```


```{r}
## ------------------------------------------------------------
## multivariate regression
## ------------------------------------------------------------
mtcars.mreg <- rfsrc(Multivar(mpg, cyl) ~., data = mtcars)
plot.variable(mtcars.mreg, m.target = "mpg", partial = TRUE, nvar = 1)
plot.variable(mtcars.mreg, m.target = "cyl", partial = TRUE, nvar = 1)
```


```{r}
## ------------------------------------------------------------
## multivariate mixed outcomes
## ------------------------------------------------------------
mtcars2 <- mtcars
mtcars2$carb <- factor(mtcars2$carb)
mtcars2$cyl <- factor(mtcars2$cyl)
mtcars.mix <- rfsrc(Multivar(carb, mpg, cyl) ~ ., data = mtcars2)
plot.variable(mtcars.mix, m.target = "cyl", target = "4", partial = TRUE, nvar = 1)
plot.variable(mtcars.mix, m.target = "cyl", target = 2, partial = TRUE, nvar = 1)

