## ``rfsrcFast`` Fast Random Forests

#### Description 

Fast approximate random forests using subsampling with forest options set to encourage computational
speed. Applies to all families.

#### Usage 
 <pre><code>
## S3 method for class 'rfsrc'
rfsrcFast(formula, data,
ntree = 500,
nsplit = 10,
bootstrap = "by.root",
ensemble = "oob",
sampsize = function(x){min(x * .632, max(150, sqrt(x)))},
samptype = "swor",
samp = NULL,
ntime = 50,
forest = FALSE,
...)
</code></pre>
#### Arguments 

* ``formula``: A symbolic description  of the model to be fit. If missing, unsupervised splitting
is implemented.
data Data frame containing the y-outcome and x-variables.
ntree Number of trees.
nsplit Non-negative integer value specifying number of random split points used to
*``split``:  a node  (deterministic splitting corresponds to the value zero and is much
slower).
* bootstrap Bootstrap protocol used in growing a tree.
* ensemble Specifies the type of ensemble. We request only out-of-sample which corresponds to "oob".
*``sampsize``: Function specifying requested size of subsampled data relative to the original
data. The requested sample size can also be passed in as a number.
*``samptype ``:   Type of bootstrap used.

*``samp``:   Bootstrap specification when "by.user" is used.
* ntime Integer value used for survival to constrain ensemble calculations to a grid of
ntime time points.
* forest Should the forest object be returned?
* ... Further arguments to be passed to rfsrc.

#### Details

Calls rfsrc under various options (including subsampling) to encourage computational speeds.
This will provide a good approximation but will not be as good as default settings of rfsrc.



#### Examples 
 ```{r}
## ------------------------------------------------------------
## Iowa housing regression example
## ------------------------------------------------------------
## load the Iowa housing data
data(housing, package = "randomForestSRC")
## do quick and *dirty* imputation
housing2 <- impute(SalePrice ~ ., housing,
ntree = 50, nimpute = 1, splitrule = "random")
## grow a fast forest
o1 <- rfsrcFast(SalePrice ~ ., housing2)
o2 <- rfsrcFast(SalePrice ~ ., housing2, nodesize = 1)
print(o1)
print(o2)
```

```{r}
## grow a fast bivariate forest
o3 <- rfsrcFast(cbind(SalePrice,Overall.Qual) ~ ., housing2)
print(o3)
```

```{r}
## ------------------------------------------------------------
## White wine classification example
## ------------------------------------------------------------
rfsrcSyn 63
data(wine, package = "randomForestSRC")
wine$quality <- factor(wine$quality)
o <- rfsrcFast(quality ~ ., wine)
print(o)
```

```{r}
## ------------------------------------------------------------
## pbc survival example
## ------------------------------------------------------------
data(pbc, package = "randomForestSRC")
o <- rfsrcFast(Surv(days, status) ~ ., pbc)
print(o)
```

```{r}
## ------------------------------------------------------------
## WIHS competing risk example
## ------------------------------------------------------------
data(wihs, package = "randomForestSRC")
o <- rfsrcFast(Surv(time, status) ~ ., wihs)
print(o)
```

#### Value
An object of class (rfsrc, grow).
Author(s)
Hemant Ishwaran and Udaya B. Kogalur
See Also
rfsrc
