
rfsrcSyn Synthetic Random Forests

#### Description 

Grows a synthetic random forest (RF) using RF machines as synthetic features. Applies only to
regression and classification settings.

#### Usage 
 <pre><code>
## S3 method for class 'rfsrc'
rfsrcSyn(formula, data, object, newdata,
ntree = 1000, mtry = NULL, nodesize = 5, nsplit = 10,
mtrySeq = NULL, nodesizeSeq = c(1:10,20,30,50,100),
min.node = 3,
fast = TRUE,
use.org.features = TRUE,
na.action = c("na.omit", "na.impute"),
oob = TRUE,
verbose = TRUE,
...)
</code></pre>
#### Arguments 

* ``formula``: A symbolic description  of the model to be fit. Must be specified unless object
is given.
64 rfsrcSyn
data Data frame containing the y-outcome and x-variables in the model. Must be
specified unless object is given.
object An object of class (rfsrc, synthetic). Not required when formula and data
are supplied.
newdata Test data used for prediction (optional).
ntree Number of trees.
mtry mtry value for over-arching synthetic forest.
nodesize Nodesize value for over-arching synthetic forest.
nsplit nsplit-randomized splitting for significantly increased speed.
mtrySeq Sequence of mtry values used for fitting the collection of RF machines. If NULL,
set to the default value p/3.
nodesizeSeq Sequence of nodesize values used for the fitting the collection of RF machines.
min.node Minimum forest averaged number of nodes a RF machine must exceed in order
to be used as a synthetic feature.
fast Use fast random forests, rfsrcFast, in place of rfsrc? Improves speed but
may be less accurate.
use.org.features
In addition to synthetic features, should the original features be used when fitting
synthetic forests?
na.action Missing value action. The default na.omit removes the entire record if even
one of its entries is NA. The action na.impute pre-imputes the data using fast
imputation via impute.rfsrc.
oob Preserve "out-of-bagness" so that error rates and VIMP are honest? Default is
yes (‘oob=TRUE’).
verbose Set to TRUE for verbose output.
... Further arguments to be passed to the rfsrc function used for fitting the synthetic
forest.
#### Details
A collection of random forests are fit using different nodesize values. The predicted values from
these machines are then used as synthetic features (called RF machines) to fit a synthetic random
forest (the original features are also used in constructing the synthetic forest). Currently only implemented
for regression and classification settings (univariate and multivariate).
Synthetic features are calculated using out-of-bag (OOB) data to avoid over-using training data.
However, to guarantee that performance values such as error rates and VIMP are honest, bootstrap
draws are fixed across all trees used in the construction of the synthetic forest and its synthetic
features. The option ‘oob=TRUE’ ensures that this happens. Change this option at your own peril.
If values for mtrySeq are given, RF machines are constructed for each combination of nodesize and
mtry values specified by nodesizeSeq mtrySeq.
rfsrcSyn 65
Value
A list with the following components:
rfMachines RF machines used to construct the synthetic features.
rfSyn The (grow) synthetic RF built over training data.
rfSynPred The predict synthetic RF built over test data (if available).
synthetic List containing the synthetic features.
opt.machine Optimal machine: RF machine with smallest OOB error rate.
Author(s)
Hemant Ishwaran and Udaya B. Kogalur
References
Ishwaran H. and Malley J.D. (2014). Synthetic learning machines. BioData Mining, 7:28.
See Also
rfsrc, rfsrcFast

#### Examples 

```{r}
## ------------------------------------------------------------
## compare synthetic forests to regular forest (classification)
## ------------------------------------------------------------
## rfsrc and rfsrcSyn calls
if (library("mlbench", logical.return = TRUE)) {
## simulate the data
ring <- data.frame(mlbench.ringnorm(250, 20))
## classification forests
ringRF <- rfsrc(classes ~., ring)
## synthetic forests
## 1 = nodesize varied
## 2 = nodesize/mtry varied
ringSyn1 <- rfsrcSyn(classes ~., ring)
ringSyn2 <- rfsrcSyn(classes ~., ring, mtrySeq = c(1, 10, 20))
## test-set performance
ring.test <- data.frame(mlbench.ringnorm(500, 20))
pred.ringRF <- predict(ringRF, newdata = ring.test)
pred.ringSyn1 <- rfsrcSyn(object = ringSyn1, newdata = ring.test)$rfSynPred
pred.ringSyn2 <- rfsrcSyn(object = ringSyn2, newdata = ring.test)$rfSynPred
66 stat.split
print(pred.ringRF)
print(pred.ringSyn1)
print(pred.ringSyn2)
}
```

```{r}
## ------------------------------------------------------------
## compare synthetic forest to regular forest (regression)
## ------------------------------------------------------------
## simulate the data
n <- 250
ntest <- 1000
N <- n + ntest
d <- 50
std <- 0.1
x <- matrix(runif(N * d, -1, 1), ncol = d)
y <- 1 * (x[,1] + x[,4]^3 + x[,9] + sin(x[,12]*x[,18]) + rnorm(n, sd = std)>.38)
dat <- data.frame(x = x, y = y)
test <- (n+1):N
## regression forests
regF <- rfsrc(y ~ ., dat[-test, ], )
pred.regF <- predict(regF, dat[test, ], importance = "none")
## synthetic forests using fast rfsrc
synF1 <- rfsrcSyn(y ~ ., dat[-test, ], fast = TRUE, newdata = dat[test, ])
synF2 <- rfsrcSyn(y ~ ., dat[-test, ], fast = TRUE,
newdata = dat[test, ], mtrySeq = c(1, 10, 20, 30, 40, 50))
## standardized MSE performance
mse <- c(tail(pred.regF$err.rate, 1),
tail(synF1$rfSynPred$err.rate, 1),
tail(synF2$rfSynPred$err.rate, 1)) / var(y[-test])
names(mse) <- c("forest", "synthetic1", "synthetic2")
print(mse)
```

```{r}
## ------------------------------------------------------------
## multivariate synthetic forests
## ------------------------------------------------------------
mtcars.new <- mtcars
mtcars.new$cyl <- factor(mtcars.new$cyl)
mtcars.new$carb <- factor(mtcars.new$carb, ordered = TRUE)
trn <- sample(1:nrow(mtcars.new), nrow(mtcars.new)/2)
mvSyn <- rfsrcSyn(cbind(carb, mpg, cyl) ~., mtcars.new[trn,])
mvSyn.pred <- rfsrcSyn(object = mvSyn, newdata = mtcars.new[-trn,])
```
