
## vimp VIMP for Single or Grouped Variables

#### Description
Calculate variable importance (VIMP) for a single variable or group of variables for training or test
data.

#### Usage
<pre><code>
## S3 method for class 'rfsrc'
vimp(object, xvar.names, m.target = NULL,
importance = c("permute", "random", "anti"), block.size = 1,
joint = FALSE, subset, seed = NULL, do.trace = FALSE, ...)
</code></pre>

#### Arguments
* object An object of class (rfsrc, grow) or (rfsrc, forest). Requires ‘forest=TRUE’
in the original rfsrc call.
* xvar.names Names of the x-variables to be used. If not specified all variables are used.
* m.target Character value for multivariate families specifying the target outcome to be
used. If left unspecified, the algorithm will choose a default target.
importance Type of VIMP.
* block.size Specifies number of trees in a block when calculating VIMP.
* joint Individual or joint VIMP?
* subset Vector indicating which rows of the grow data to restrict VIMP calculations to;
i.e. this option yields VIMP which is restricted to a specific subset of the data.
Note that the vector should correspond to the rows of object$xvar and not the original data passed in the grow call. All rows used if not specified.
* seed Negative integer specifying seed for the random number generator.
* do.trace Number of seconds between updates to the user on approximate time to completion.
* ... Further arguments passed to or from other methods.

#### Details
Using a previously grown forest, calculate the VIMP for variables xvar.names. By default, VIMP
is calculated for the original data, but the user can specify a new test data for the VIMP calculation
using newdata. See rfsrc for more details about how VIMP is calculated.
Joint VIMP is requested using ‘joint’ and equals importance for a group of variables when the
group is perturbed simultaneously.

#### Value
An object of class (rfsrc, predict) containing importance values.

#### Author(s)
Hemant Ishwaran and Udaya B. Kogalur

#### References
Ishwaran H. (2007). Variable importance in binary regression trees and forests, Electronic J. Statist.,
1:519-537.
See Also
rfsrc

```{r}
### Examples

```{r}
## ------------------------------------------------------------
## classification example
## showcase different vimp
## ------------------------------------------------------------
iris.obj <- rfsrc(Species ~ ., data = iris)
# Permutation vimp
print(vimp(iris.obj)$importance)
# Random daughter vimp
print(vimp(iris.obj, importance = "random")$importance)
# Joint permutation vimp
print(vimp(iris.obj, joint = TRUE)$importance)
# Paired vimp
print(vimp(iris.obj, c("Petal.Length", "Petal.Width"), joint = TRUE)$importance)
print(vimp(iris.obj, c("Sepal.Length", "Petal.Width"), joint = TRUE)$importance)
```

```{r}
## ------------------------------------------------------------
## regression example
## ------------------------------------------------------------
airq.obj <- rfsrc(Ozone ~ ., airquality)
print(vimp(airq.obj))
## ------------------------------------------------------------
## regression example where vimp is calculated on test data
## ------------------------------------------------------------
set.seed(100080)
train <- sample(1:nrow(airquality), size = 80)
airq.obj <- rfsrc(Ozone~., airquality[train, ])
wihs 83
#training data vimp
print(airq.obj$importance)
print(vimp(airq.obj)$importance)
#test data vimp
print(vimp(airq.obj, newdata = airquality[-train, ])$importance)
```

```{r}
## ------------------------------------------------------------
## survival example
## study how vimp depends on tree imputation
## makes use of the subset option
## ------------------------------------------------------------
data(pbc, package = "randomForestSRC")
# determine which records have missing values
which.na <- apply(pbc, 1, function(x){any(is.na(x))})
# impute the data using na.action = "na.impute"
pbc.obj <- rfsrc(Surv(days,status) ~ ., pbc, nsplit = 3,
na.action = "na.impute", nimpute = 1)
# compare vimp based on records with no missing values
# to those that have missing values
# note the option na.action="na.impute" in the vimp() call
vimp.not.na <- vimp(pbc.obj, subset = !which.na, na.action = "na.impute")$importance
vimp.na <- vimp(pbc.obj, subset = which.na, na.action = "na.impute")$importance
print(data.frame(vimp.not.na, vimp.na))
```
